%!TEX root = ../Thesis.tex
\chapter*{\englishproblemname}
\addcontentsline{toc}{chapter}{\englishproblemname}

\textbf{Introduction: }
This project aims to develop a method for multi-agent control of Autonomous Underwater Vehicles (AUVs) by combining the 3D hand position concept with a selected multi-agent control algorithm. The goal is to enable accurate formation control, path following, and collision avoidance for AUVs in a multi-agent system.

\textbf{Objective: }
\begin{itemize}
    \item Perform a literature study on control methods for multi-agent systems with double integrator dynamics, including relevant papers on the 3D hand position concept and multi-agent control algorithms.
    \item Select a suitable multi-agent control algorithm and integrate it with the 3D hand position concept to develop a control strategy for multi-agent control of AUVs.
    \item Develop the theory for the combined control method, including stability analysis.
    \item Conduct simulation studies to validate the effectiveness and performance of the developed control method in formation control, path following, and collision avoidance tasks.
\end{itemize}


\textbf{Expected Outcome: }
The expected outcome of this project is an advanced control method that combines the 3D hand position concept with a selected multi-agent control algorithm to achieve accurate and robust multi-agent control of AUVs. The developed control method will be validated through simulation studies, demonstrating its effectiveness in various scenarios.


\textbf{Conclusion: }
By combining the 3D hand position concept with a selected multi-agent control algorithm, this project aims to provide a novel control method for AUVs in a multi-agent system. The proposed method has the potential to significantly enhance the formation control, path following, and collision avoidance capabilities of AUVs, contributing to advancements in autonomous underwater robotics.


% \textcolor{DarkOrange}{To be formulated thoroughly later.}

% The parts I have addressed so far are marked in bold. I have marked the general multi-agent literature study in italics, as it was partly covered while writing the conference paper.
% \begin{itemize}
%     \item Perform a literature study on control methods of multi-agent systems with double integrator dynamics.
%     \item \textbf{Choose a multi-agent control method that is suitable to combine with the 3D hand position method and merge these, to develop a control method for multi-agent control of AUVs.}
%     \item \textbf{Validate the theoretical results through a simulation study.}
%     \item Validate the theoretical results through high-fidelity simulations/experiments.
%     \item \textit{Do a literature study on multi-agent control methods for marine vehicles (ASVs and AUVs), to put your results into its research context.}
% \end{itemize}
%
\clearpage
